---
title: "scientific programming project"
output: html_document
---

# Scientific Programming Project: A. Eliot Hall

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

### Setup and Importing:

Libraries:

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(tibble)
library(data.table)
library(mapview)
library(sf)
library(leaflet)
library(leafsync)
library(maps)
library(rstudioapi)

```

Setting Directory and Import Bird Data

```{r}
setwd(dirname(getActiveDocumentContext()$path))

Bird_Species <- read.csv("Bird_Species.csv")
Bird_Trait <- read.csv("Bird_Trait.csv")
Bird_Sites <- read.csv("Bird_Sites.csv")
```

Setting up data-frames so column naming is consistent and only relevant variables are present

```{r}
colnames(Bird_Species)[1] <- "City_ID"
 
Bird_Trait <- Bird_Trait[, c(-2, -3 , -10:-19)] 
Bird_Sites <- Bird_Sites[,-5:-8]
```

Import Amphibian Data

```{r}
Amphibian_Species <- read.csv("Amphibian_Species.csv")
Amphibian_Trait <- read.csv("Amphibian_Trait.csv")
Amphibian_Sites <- read.csv("Amphibian_Site.csv")
```

Setting up data-frames so column naming is consistent and only relevant variables are present

```{r}
colnames(Amphibian_Species)[1] <- "SiteID"
 
Amphibian_Trait <- Amphibian_Trait[, c(-2, -4, -6)] 
Amphibian_Sites <- Amphibian_Sites[,-5:-12]
```

### Pre-processing / Cleaning:

Merge Species + Sites Data-frames by SiteID / City_ID:

```{r}
Amphibian_Species_Sites <- left_join(Amphibian_Sites, Amphibian_Species, by = "SiteID")
Bird_Species_Sites <- left_join(Bird_Sites, Bird_Species, by = "City_ID")
```

Check for NA's and fix:

```{r}
sum(is.na(Bird_Species_Sites))
sum(is.na(Amphibian_Species_Sites))
# join function automatically gets rid of NAs :) yay!

# which(is.na(Amphibian_Trait)) # no NAs :)

# which(is.na(Bird_Trait)) #many NAs....

# creating a separate dataframe for the species with missing values to keep track 
Bird_Trait_Missing <- Bird_Trait %>% filter(is.na(Bird_Trait$Hand.wing.index) | is.na(Bird_Trait$Fruit.nectar) | is.na(Bird_Trait$Invertebrate) | is.na(Bird_Trait$Omnivore) | is.na(Bird_Trait$Plant.seed) | is.na(Bird_Trait$Vertebrates.scavenger))

# removing NAs from main dataframe
Bird_Trait <- na.omit(Bird_Trait)

# double check all dataframes for NAs again
sum(is.na(Bird_Species_Sites))
sum(is.na(Bird_Trait))
sum(is.na(Amphibian_Species_Sites))
sum(is.na(Amphibian_Trait))
```

Fixing Inconsistent Labeling:

```{r}
# fixing problem found in labeling:
#  Amphibian movement.distances has "Reduced" and "Reduce" :( and Amphibian diet.breadth has "Generalist" and "Generalists"
Amphibian_Trait$Movement.Distances[Amphibian_Trait$Movement.Distances == "Reduce"] <- "Reduced"
Amphibian_Trait$Diet.breadth[Amphibian_Trait$Diet.breadth == "Generalists"] <- "Generalist"
Amphibian_Trait$Diet.breadth[Amphibian_Trait$Diet.breadth == "Specialists"] <- "Specialist"
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

## Creating Urban Syndromes - Birds

#### Group Generalists and Specialists in Bird_Trait Data-frame:

```{r}
# first convert relevant columns to character so that replacing name is easier
Bird_Trait$Fruit.nectar <- as.character(Bird_Trait$Fruit.nectar)
Bird_Trait$Invertebrate <- as.character(Bird_Trait$Invertebrate)
Bird_Trait$Omnivore <- as.character(Bird_Trait$Omnivore)
Bird_Trait$Plant.seed <- as.character(Bird_Trait$Plant.seed)
Bird_Trait$Vertebrates.scavenger <- as.character(Bird_Trait$Vertebrates.scavenger)

# Replace Binary Values of Diet with Generalist or Specialist
Bird_Trait$Fruit.nectar[Bird_Trait$Fruit.nectar == "1"] <- "Specialist"
Bird_Trait$Invertebrate[Bird_Trait$Invertebrate == "1"] <- "Specialist"
Bird_Trait$Plant.seed[Bird_Trait$Plant.seed == "1"] <- "Specialist"

Bird_Trait$Omnivore[Bird_Trait$Omnivore == "1"] <- "Generalist" 
Bird_Trait$Vertebrates.scavenger[Bird_Trait$Vertebrates.scavenger == "1"] <- "Generalist"

Bird_Trait <- Bird_Trait %>% unite("Diet.breadth.S", c(Fruit.nectar, Invertebrate, Plant.seed)) #combining specialists
Bird_Trait <- Bird_Trait %>% unite("Diet.breath.G", c(Omnivore, Vertebrates.scavenger)) #combining generalist #even though I dont technically need to do this because I won't be using this column, it allows me to easily check to make sure that I combined Specialist properly and there is no incorrect overlap 

#unique(Bird_Trait$Diet.breadth.S) #checking string names in combined column to make sure I rename all of them correctly
Bird_Trait$Diet.breadth.S[Bird_Trait$Diet.breadth.S == "0_0_0"] <- "Generalist"  # 0_0_0 is a generalist because if they are not a specialist then they are a generalist
Bird_Trait$Diet.breadth.S[Bird_Trait$Diet.breadth.S == "0_Specialist_0"] <- "Specialist"
Bird_Trait$Diet.breadth.S[Bird_Trait$Diet.breadth.S == "Specialist_0_0"] <- "Specialist"
Bird_Trait$Diet.breadth.S[Bird_Trait$Diet.breadth.S == "0_0_Specialist"] <- "Specialist"

#finalizing dataframe
colnames(Bird_Trait)[3] <- "Diet.breadth" #renaming Diet column now that it contains all combined information for specialist and generalist 
Bird_Trait <- Bird_Trait[, -4] #removing excess column
```

#### Looking at Density Plot Hand.wing.index values - determine best way to bin them:

```{r}
probs <- c(0.25, 0.5, 0.75)
quantiles <- quantile(Bird_Trait$Hand.wing.index, prob=probs)
quantiles
mean(Bird_Trait$Hand.wing.index)

density_plot <- ggplot(Bird_Trait, aes(x=Hand.wing.index)) +
  geom_density(color="#3399CC", fill="#99CCCC") + 
  geom_vline(aes(xintercept = mean(Hand.wing.index)), color="blue", linetype="dashed", size=.75) +
  geom_vline(aes(xintercept=18.07887), color="#6600FF",  linetype="dashed") +
   geom_vline(aes(xintercept=25.57697), color="#6600FF",  linetype="dashed") +
  geom_vline(aes(xintercept=38.65697), color="#6600FF",  linetype="dashed") +
  annotate("text", x = 16.5, y = 0.00575, label = "25th Quartile", angle = 90)  +
  annotate("text", x = 24, y = 0.005, label = "Median", angle = 90) +
  annotate("text", x = 37,  y = 0.00575, label = "75th Quartile", angle = 90)  +
  annotate("text", x = 28, y = 0.005, label = "Mean", angle = 90, color = "blue") +
  labs(x = "Hand-Wing Index", y = "Density", title = "Density Plot of Hand-wing Index Measure") + theme(plot.title = element_text(hjust = 0.5), axis.line=element_line(color="black")) + scale_x_continuous(expand=c(0, 0), limits=c(0, 90)) + scale_y_continuous(expand=c(0, 0), limits=c(0, 0.04))

density_plot
```

#### Grouping Reduced, Moderate, and High Mobility Categorizes in Bird_Trait Data-frame:

```{r}

Bird_Trait <- within(Bird_Trait, {
  Movement.Distances <- NA
  Movement.Distances[Hand.wing.index < 18.07887] <- "Reduced"
  Movement.Distances[Hand.wing.index >= 18.07887 & Hand.wing.index < 38.65697] <- "Moderate"
  Movement.Distances[Hand.wing.index >= 38.65697] <- "High"
} )


```

#### Creating Syndromes:

```{r}
# Mobile Generalist: Moderate to High Mobility & Generalist 
# Mobile Specialist: Moderate to High Mobility & Specialist
# Central Forager: Low to Moderate Mobility & Generalist 
# Site Specialist: Low to Moderate Mobility & Specialist

#for birds the categorization is not a problem because I can still split High and Low mobility according to the median Hand.wing.index 

Bird_Trait <- within(Bird_Trait, {
  Syndrome <- NA
  Syndrome[Diet.breadth == "Generalist" & Hand.wing.index >= 25.57697] <- "M.G."
  Syndrome[Diet.breadth == "Generalist" & Hand.wing.index < 25.57697] <- "C.F"
  Syndrome[Diet.breadth == "Specialist" & Hand.wing.index >= 25.57697] <- "M.S."
  Syndrome[Diet.breadth == "Specialist" & Hand.wing.index < 25.57697] <- "S.S"
})

sum(is.na(Bird_Trait))
```

## Creating Urban Syndromes - Amphibians

```{r}
# Mobile Generalist: Moderate to High Mobility & Generalist 
# Mobile Specialist: Moderate to High Mobility & Specialist
# Central Forager: Low to Moderate Mobility & Generalist 
# Site Specialist: Low to Moderate Mobility & Specialist

# found problem: 
# the "Moderate" mobility category is messing up categorizing the syndromes - will have to split the Moderate 
# Merge Moderate Mobility Generalist with High Mobility Generalist (Mobile Generalist) and Moderate Mobility Specialist with Reduced Mobility Specialist (Site Specialist) 

Amphibian_Trait <- within(Amphibian_Trait, {
  Syndrome <- NA
  Syndrome[Diet.breadth == "Generalist" & Movement.Distances == "High"] <- "M.G."
  Syndrome[Diet.breadth == "Generalist" & Movement.Distances == "Moderate"] <- "M.G."
  Syndrome[Diet.breadth == "Generalist" & Movement.Distances == "Reduced"] <- "C.F."
  Syndrome[Diet.breadth == "Specialist" & Movement.Distances == "High"] <- "M.S."
  Syndrome[Diet.breadth == "Specialist" & Movement.Distances == "Reduced"] <- "S.S."
  Syndrome[Diet.breadth == "Specialist" & Movement.Distances == "Moderate"] <- "S.S."
})

sum(is.na(Amphibian_Trait))
```

## Creating Occurrence Data per Syndrome for Spatial Distribution Mapping - Birds

#### Converting Occurrence Data to be Syndrome Specific: Birds

```{r}
#creating occurrence column for each syndrome (if S.S = 1, else = 0 ...... etc.)
Bird_Trait <- Bird_Trait %>% mutate(S.S = if_else(Syndrome == "S.S", 1, 0))
Bird_Trait <- Bird_Trait %>% mutate(M.S = if_else(Syndrome == "M.S.", 1, 0))
Bird_Trait <- Bird_Trait %>% mutate(M.G = if_else(Syndrome == "M.G.", 1, 0))
Bird_Trait <- Bird_Trait %>% mutate(C.F = if_else(Syndrome == "C.F", 1, 0))

# checking to make sure the occurrences have been inputted properly - the number of each Syndrome should make the sum of the occurrences  
sum(Bird_Trait$Syndrome == "S.S") == sum(Bird_Trait$S.S)
sum(Bird_Trait$Syndrome == "M.S.") == sum(Bird_Trait$M.S)
sum(Bird_Trait$Syndrome == "M.G.") == sum(Bird_Trait$M.G)
sum(Bird_Trait$Syndrome == "C.F") == sum(Bird_Trait$C.F)
```

```{r}
#convert species name from "X.y" to "X y" structure (because I will need to match by species name)
Bird_Trait$X <- lapply(Bird_Trait$X, gsub, pattern = " ", replacement = ".", fixed = TRUE)
```

```{r}
# want to create a Bird_Species_Sites df per syndrome - this way, it will still be species occurrence - but the species will be "filtered" by their syndrome
#----------------------------------------------------------------------------------

# Site Specialist (S.S): 

# separating syndrome and structuring appropriately 
S.S_Bird_1 <- Bird_Trait[, c(1, 6)]
rownames(S.S_Bird_1) <- NULL
S.S_Bird_1 <- column_to_rownames(S.S_Bird_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
S.S_Bird_1 <- S.S_Bird_1 %>% filter(S.S == 1)

# transposing so that I can match by column name (Species)
S.S_Bird <- transpose(S.S_Bird_1)
rownames(S.S_Bird) <- colnames(S.S_Bird_1)
colnames(S.S_Bird) <- rownames(S.S_Bird_1)

# structuring Bird_Species df appropriately so I can match by column name (species)
Bird_Species <- column_to_rownames(Bird_Species, var = "City_ID")

# Creating S.S Syndrome specific occurrence df based on matching Species names
Bird_Species_S.S <- Bird_Species %>% select(matches(colnames(S.S_Bird)))

#---------------------------------------------------------------------------------

# Mobile Specialist(M.S):

# separating syndrome and structuring appropriately 
M.S_Bird_1 <- Bird_Trait[, c(1, 7)]
rownames(M.S_Bird_1) <- NULL
M.S_Bird_1 <- column_to_rownames(M.S_Bird_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
M.S_Bird_1 <- M.S_Bird_1 %>% filter(M.S == 1)

# transposing so that I can match by column name (Species)
M.S_Bird <- transpose(M.S_Bird_1)
rownames(M.S_Bird) <- colnames(M.S_Bird_1)
colnames(M.S_Bird) <- rownames(M.S_Bird_1)

# Creating M.S Syndrome specific occurrence df based on matching Species names
Bird_Species_M.S <- Bird_Species %>% select(matches(colnames(M.S_Bird)))

#----------------------------------------------------------------------------------

# Mobile Generalist (M.G):

# separating syndrome and structuring appropriately 
M.G_Bird_1 <- Bird_Trait[, c(1, 8)]
rownames(M.G_Bird_1) <- NULL
M.G_Bird_1 <- column_to_rownames(M.G_Bird_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
M.G_Bird_1 <- M.G_Bird_1 %>% filter(M.G == 1)

# transposing so that I can match by column name (Species)
M.G_Bird <- transpose(M.G_Bird_1)
rownames(M.G_Bird) <- colnames(M.G_Bird_1)
colnames(M.G_Bird) <- rownames(M.G_Bird_1)

# Creating M.G Syndrome specific occurrence df based on matching Species names
Bird_Species_M.G <- Bird_Species %>% select(matches(colnames(M.G_Bird)))

#----------------------------------------------------------------------------------

# Central Forager (C.F):

# separating syndrome and structuring appropriately 
C.F_Bird_1 <- Bird_Trait[, c(1, 9)]
rownames(C.F_Bird_1) <- NULL
C.F_Bird_1 <- column_to_rownames(C.F_Bird_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
C.F_Bird_1 <- C.F_Bird_1 %>% filter(C.F == 1)

# transposing so that I can match by column name (Species)
C.F_Bird <- transpose(C.F_Bird_1)
rownames(C.F_Bird) <- colnames(C.F_Bird_1)
colnames(C.F_Bird) <- rownames(C.F_Bird_1)

# Creating C.F Syndrome specific occurrence df based on matching Species names
Bird_Species_C.F <- Bird_Species %>% select(matches(colnames(C.F_Bird)))
```

```{r}
# isolating locations in which syndromes occur in order to match relevant coordinates
#----------------------------------------------------------------------------------

# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Bird_Species_S.S$sum <- rowSums(Bird_Species_S.S)
Bird_Species_S.S <- Bird_Species_S.S[-which(Bird_Species_S.S$sum == 0), ]
# checking that all 0's were removed 
which(Bird_Species_S.S$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Bird_Species_M.S$sum <- rowSums(Bird_Species_M.S)
Bird_Species_M.S <- Bird_Species_M.S[-which(Bird_Species_M.S$sum == 0), ]
# checking that all 0's were removed 
which(Bird_Species_M.S$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Bird_Species_M.G$sum <- rowSums(Bird_Species_M.G)
Bird_Species_M.G <- Bird_Species_M.G[-which(Bird_Species_M.G$sum == 0), ]
# checking that all 0's were removed 
which(Bird_Species_M.G$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Bird_Species_C.F$sum <- rowSums(Bird_Species_C.F)
Bird_Species_C.F <- Bird_Species_C.F[-which(Bird_Species_C.F$sum == 0), ]
# checking that all 0's were removed 
which(Bird_Species_C.F$sum ==0) 
```

## Mapping Spatial Distribution of Urban Syndromes - Birds

#### Mapping Spatial Distribution of Urban Syndromes: Birds

```{r}
# altering rownames so I to merge dataframes easier
Bird_Sites <- column_to_rownames(Bird_Sites, var = "City_ID")
```

```{r}
# creating df of syndrome locations  (lat, lng, and location data)
    # S.S:
S.S_location <- merge(Bird_Sites, Bird_Species_S.S, by = 'row.names', all = FALSE)
S.S_location <- S.S_location[, c(-5:-1217)]
colnames(S.S_location)[1] <- "City_ID"
    # M.S: 
M.S_location <- merge(Bird_Sites, Bird_Species_M.S, by = 'row.names', all = FALSE)
M.S_location <- M.S_location[, c(-5:-1183)]
colnames(M.S_location)[1] <- "City_ID"
    # M.G: 
M.G_location <- merge(Bird_Sites, Bird_Species_M.G, by = 'row.names', all = FALSE)
M.G_location <- M.G_location[, c(-5:-470)]
colnames(M.G_location)[1] <- "City_ID"
    # C.F: 
C.F_location <- merge(Bird_Sites, Bird_Species_C.F, by = 'row.names', all = FALSE)
C.F_location <- C.F_location[, c(-5:-438)]
colnames(C.F_location)[1] <- "City_ID"


# converting to appropriate form in order to plot, identifying coordinates and map system type for each syndrome df
test= st_as_sf(S.S_location, coords = c("longitude", "latitude"), crs = 4326)
test_1= st_as_sf(M.S_location, coords = c("longitude", "latitude"), crs = 4326)
test_2= st_as_sf(M.G_location, coords = c("longitude", "latitude"), crs = 4326)
test_3= st_as_sf(C.F_location, coords = c("longitude", "latitude"), crs = 4326)


# in order to focus/visualize a more specific area, want to narrow to North & South America 
western_hemisphere_data <- list(
  SiteSpecialist = st_as_sf(S.S_location[S.S_location$longitude < -30 & S.S_location$latitude > -30, ], coords = c("longitude", "latitude"), crs = 4326),
  MobileSpecialist = st_as_sf(M.S_location[M.S_location$longitude < -30 & M.S_location$latitude > -30, ], coords = c("longitude", "latitude"), crs = 4326),
  MobileGeneralist = st_as_sf(M.G_location[M.G_location$longitude < -30 & M.G_location$latitude > -30, ], coords = c("longitude", "latitude"), crs = 4326),
  CentralForager = st_as_sf(C.F_location[C.F_location$longitude < -30 & C.F_location$latitude > -30, ], coords = c("longitude", "latitude"), crs = 4326)
)

# syndrome names for the layers
layer_names <- c("Site Specialist", "Mobile Specialist", "Mobile Generalist", "Central Forager")

# Maps
wholemap <- mapview(list(test, test_1, test_2, test_3), layer.name = layer_names, col.regions = c("#40dde9", "#fe7191", "#00bf0b", "#f9d12e"), alpha = 0.35, cex = 3) 

westernmap <- mapview(western_hemisphere_data, layer.name = layer_names, col.regions = c("#40dde9", "#fe7191", "#00bf0b", "#f9d12e"), alpha = 0.35, cex = 3)

latticeView(wholemap, westernmap) 

```

#### Mapping the Spatial Distribution of Urban Syndromes by Site Counts, rather than just Occurrence

-   I am doing this because a lot of the cities overlap (reflected by the coordinates) but not necessarily the specific sites (reflected by Site_ID)

```{r}
  #S.S:
t <- test[,-1]
S.S.1 <- t %>% group_by(geometry) %>%
  summarise( count = n() ) %>% ungroup()
sst1 <- st_join(S.S.1, t)
sst1 <- distinct(sst1, geometry, .keep_all = TRUE)
  #M.S
t1 <- test_1[,-1]
M.S.1 <- t1 %>% group_by(geometry) %>%
  summarise( count = n() ) %>% ungroup()
mst1 <- st_join(M.S.1, t1)
mst1 <- distinct(mst1, geometry, .keep_all = TRUE)
  #M.G
t2 <- test_2[,-1]
M.G.1 <- t2 %>% group_by(geometry) %>%
  summarise( count = n() ) %>% ungroup()
mgt1 <- st_join(M.G.1, t2)
mgt1 <- distinct(mgt1, geometry, .keep_all = TRUE)
  #C.F
t3 <- test_3[,-1]
C.F.1 <- t3 %>% group_by(geometry) %>%
  summarise( count = n() ) %>% ungroup()
cft1 <- st_join(C.F.1, t3)
cft1 <- distinct(cft1, geometry, .keep_all = TRUE)


# mapview showing count (size of node) - per syndrome
wholecountmap <- mapview(list(sst1,mst1,mgt1,cft1), layer.name = layer_names, col.regions = c("#40dde9", "#fe7191", "#00bf0b", "#f9d12e"), alpha = 0.35, cex = "count")
wholecountmap

```

## Number of Species per Syndrome - Birds

```{r}
# creating a dataframe of the sum of species for each Syndrome, to visualize the proportion of Bird Species that belong to each Syndrome
#---------------------------------------------------------------------------------

#creating species counts per syndrome df:
species_counts <- data.frame("Site Specialist" = colSums(S.S_Bird_1), "Mobile Generalist" = colSums(M.G_Bird_1), "Mobile Specialist" = colSums(M.S_Bird_1), "Central Forager" = colSums(C.F_Bird_1))
rownames(species_counts) <- "Species_Count"

#structing the df appropiately so that it is easier to use in ggplot2:
species_counts_plot <- transpose(species_counts)
rownames(species_counts_plot) <- colnames(species_counts)
colnames(species_counts_plot) <- rownames(species_counts)
species_counts_plot <- rownames_to_column(species_counts_plot, var = "Syndromes")

#barchart showing the number of Bird Species per Syndrome
species_count_per_syndrome <- ggplot(species_counts_plot, aes(x = Syndromes, y = Species_Count, fill = Syndromes)) + geom_bar(stat = "identity", color = "black") + theme_minimal() + scale_fill_manual(values=c("#f9d12e", "#00bf0b", "#fe7191", "#40dde9")) + labs(title = "Bird Species Count per Syndrome", y = "Number of Species") + theme(plot.title = element_text(hjust = 0.5), axis.line=element_line(color="black")) + scale_x_discrete(labels=c("Central.Forager" = "Central Forager", "Mobile.Generalist" = "Mobile Generalist", "Mobile.Specialist" = "Mobile Specialist", "Site.Specialist" = "Site Specialist")) + scale_y_continuous(expand=c(0, 0), limits=c(0, 1250)) + theme(legend.position = "none") 


species_count_per_syndrome

```

## Creating Occurrence Data per Syndrome for Spatial Distribution Mapping - Amphibian

#### Converting Occurrence Data to be Syndrome Specific: Amphibian

```{r}
#creating occurrence column for each syndrome (if S.S = 1, else = 0 ...... etc.)
Amphibian_Trait <- Amphibian_Trait %>% mutate(S.S = if_else(Syndrome == "S.S.", 1, 0))
Amphibian_Trait <- Amphibian_Trait %>% mutate(M.S = if_else(Syndrome == "M.S.", 1, 0))
Amphibian_Trait <- Amphibian_Trait %>% mutate(M.G = if_else(Syndrome == "M.G.", 1, 0))
Amphibian_Trait <- Amphibian_Trait %>% mutate(C.F = if_else(Syndrome == "C.F.", 1, 0))

# checking to make sure the occurrences have been inputted properly - the number of each Syndrome should make the sum of the occurrences  
sum(Amphibian_Trait$Syndrome == "S.S.") == sum(Amphibian_Trait$S.S)
sum(Amphibian_Trait$Syndrome == "M.S.") == sum(Amphibian_Trait$M.S)
sum(Amphibian_Trait$Syndrome == "M.G.") == sum(Amphibian_Trait$M.G)
sum(Amphibian_Trait$Syndrome == "C.F.") == sum(Amphibian_Trait$C.F)

```

```{r}
#convert species name from "X y" to "X.y" structure (because I will need to match by species name)
Amphibian_Trait$X <- lapply(Amphibian_Trait$X, gsub, pattern = " ", replacement = ".", fixed = TRUE)
```

```{r}
# want to create a Amphibian_Species_Sites df per syndrome - this way, it will still be species occurrence - but the species will be "filtered" by their syndrome
#----------------------------------------------------------------------------------

# Site Specialist (S.S): 

# separating syndrome and structuring appropriately 
S.S_Amphibian_1 <- Amphibian_Trait[, c(1, 5)]
rownames(S.S_Amphibian_1) <- NULL
S.S_Amphibian_1 <- column_to_rownames(S.S_Amphibian_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
S.S_Amphibian_1 <- S.S_Amphibian_1 %>% filter(S.S == 1)

# transposing so that I can match by column name (Species)
S.S_Amphibian <- transpose(S.S_Amphibian_1)
rownames(S.S_Amphibian) <- colnames(S.S_Amphibian_1)
colnames(S.S_Amphibian) <- rownames(S.S_Amphibian_1)

# structuring Bird_Species df appropriately so I can match by column name (species)
Amphibian_Species <- column_to_rownames(Amphibian_Species, var = "SiteID")

# Creating S.S Syndrome specific occurrence df based on matching Species names
Amphibian_Species_S.S <- Amphibian_Species %>% select(matches(colnames(S.S_Amphibian)))

#---------------------------------------------------------------------------------

# Mobile Specialist(M.S):

# separating syndrome and structuring appropriately 
M.S_Amphibian_1 <- Amphibian_Trait[, c(1, 6)]
rownames(M.S_Amphibian_1) <- NULL
M.S_Amphibian_1 <- column_to_rownames(M.S_Amphibian_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
M.S_Amphibian_1 <- M.S_Amphibian_1 %>% filter(M.S == 1)

# transposing so that I can match by column name (Species)
M.S_Amphibian <- transpose(M.S_Amphibian_1)
rownames(M.S_Amphibian) <- colnames(M.S_Amphibian_1)
colnames(M.S_Amphibian) <- rownames(M.S_Amphibian_1)

# Creating M.S Syndrome specific occurrence df based on matching Species names
Amphibian_Species_M.S <- Amphibian_Species %>% select(matches(colnames(M.S_Amphibian)))

#----------------------------------------------------------------------------------

# Mobile Generalist (M.G):

# separating syndrome and structuring appropriately 
M.G_Amphibian_1 <- Amphibian_Trait[, c(1, 7)]
rownames(M.G_Amphibian_1) <- NULL
M.G_Amphibian_1 <- column_to_rownames(M.G_Amphibian_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
M.G_Amphibian_1 <- M.G_Amphibian_1 %>% filter(M.G == 1)

# transposing so that I can match by column name (Species)
M.G_Amphibian <- transpose(M.G_Amphibian_1)
rownames(M.G_Amphibian) <- colnames(M.G_Amphibian_1)
colnames(M.G_Amphibian) <- rownames(M.G_Amphibian_1)

# Creating M.G Syndrome specific occurrence df based on matching Species names
Amphibian_Species_M.G <- Amphibian_Species %>% select(matches(colnames(M.G_Amphibian)))

#----------------------------------------------------------------------------------

# Central Forager (C.F):

# separating syndrome and structuring appropriately 
C.F_Amphibian_1 <- Amphibian_Trait[, c(1, 8)]
rownames(C.F_Amphibian_1) <- NULL
C.F_Amphibian_1 <- column_to_rownames(C.F_Amphibian_1, var = "X")

# removing any species that DO NOT fall within this syndrome (indicated by 0) - since I want this df to be specific to this syndrome
C.F_Amphibian_1 <- C.F_Amphibian_1 %>% filter(C.F == 1)

# transposing so that I can match by column name (Species)
C.F_Amphibian <- transpose(C.F_Amphibian_1)
rownames(C.F_Amphibian) <- colnames(C.F_Amphibian_1)
colnames(C.F_Amphibian) <- rownames(C.F_Amphibian_1)

# Creating C.F Syndrome specific occurrence df based on matching Species names
Amphibian_Species_C.F <- Amphibian_Species %>% select(matches(colnames(C.F_Amphibian)))
```

```{r}
# isolating locations in which syndromes occur in order to match relevant coordinates
#----------------------------------------------------------------------------------

# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Amphibian_Species_S.S$sum <- rowSums(Amphibian_Species_S.S)
Amphibian_Species_S.S <- Amphibian_Species_S.S[-which(Amphibian_Species_S.S$sum == 0), ]
# checking that all 0's were removed 
which(Amphibian_Species_S.S$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Amphibian_Species_M.S$sum <- rowSums(Amphibian_Species_M.S)
Amphibian_Species_M.S <- Amphibian_Species_M.S[-which(Amphibian_Species_M.S$sum == 0), ]
# checking that all 0's were removed 
which(Amphibian_Species_M.S$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Amphibian_Species_M.G$sum <- rowSums(Amphibian_Species_M.G)
Amphibian_Species_M.G <- Amphibian_Species_M.G[-which(Amphibian_Species_M.G$sum == 0), ]
# checking that all 0's were removed 
which(Amphibian_Species_M.G$sum ==0)


# removing sites where there are no occurrences (sum of all species occurrence - sum across rows - is 0)
Amphibian_Species_C.F$sum <- rowSums(Amphibian_Species_C.F)
Amphibian_Species_C.F <- Amphibian_Species_C.F[-which(Amphibian_Species_C.F$sum == 0), ]
# checking that all 0's were removed 
which(Amphibian_Species_C.F$sum ==0)
```

## Mapping Spatial Distribution of Urban Syndromes - Amphibians

#### Mapping Spatial Distribution of Urban Syndromes: Amphibians

```{r}
# altering rownames so I to merge dataframes easier
Amphibian_Sites <- column_to_rownames(Amphibian_Sites, var = "SiteID")
```

```{r}
# creating df of syndrome locations  (lat, lng, and location data)
    # S.S:
S.S_location_amph <- merge(Amphibian_Sites, Amphibian_Species_S.S, by = 'row.names', all = FALSE)
S.S_location_amph <- S.S_location_amph[, c(-5:-28)]
colnames(S.S_location_amph)[1] <- "City_ID"
    # M.S: 
M.S_location_amph <- merge(Amphibian_Sites, Amphibian_Species_M.S, by = 'row.names', all = FALSE)
M.S_location_amph <- M.S_location_amph[, c(-5:-8)]
colnames(M.S_location_amph)[1] <- "City_ID"
    # M.G: 
M.G_location_amph <- merge(Amphibian_Sites, Amphibian_Species_M.G, by = 'row.names', all = FALSE)
M.G_location_amph <- M.G_location_amph[, c(-5:-78)]
colnames(M.G_location_amph)[1] <- "City_ID"
    # C.F: 
C.F_location_amph <- merge(Amphibian_Sites, Amphibian_Species_C.F, by = 'row.names', all = FALSE)
C.F_location_amph <- C.F_location_amph[, c(-5:-46)]
colnames(C.F_location_amph)[1] <- "City_ID"


# converting to appropriate form in order to plot, identifying coordinates and map system type for each syndrome df
amph_test <-  st_as_sf(S.S_location_amph, coords = c("Longitude", "Latitude"), crs = 4326)
amph_test <- amph_test[,c(-1,-2)]
amph_test_1 <- st_as_sf(M.S_location_amph, coords = c("Longitude", "Latitude"), crs = 4326)
amph_test_1 <- amph_test_1[,c(-1,-2)]
amph_test_2 <- st_as_sf(M.G_location_amph, coords = c("Longitude", "Latitude"), crs = 4326)
amph_test_2 <- amph_test_2[,c(-1,-2)]
amph_test_3 <-  st_as_sf(C.F_location_amph, coords = c("Longitude", "Latitude"), crs = 4326)
amph_test_3 <- amph_test_3[,c(-1,-2)]

amph_list <- list(amph_test, amph_test_1, amph_test_2, amph_test_3)


# syndrome names for the layers
layer_names <- c("Site Specialist", "Mobile Specialist", "Mobile Generalist", "Central Forager")

# Create a map
amphmap <- mapview(x = amph_list, layer.name = layer_names, col.regions = c("#40dde9", "#fe7191", "#00bf0b", "#f9d12e"), alpha = 0.35, cex = 3)

amphmap
```

## Number of Species per Syndrome - Amphibian

```{r}
# creating a dataframe of the sum of species for each Syndrome, to visualize the proportion of Amphibian Species that belong to each Syndrome
#---------------------------------------------------------------------------------



#creating species counts per syndrome df:
amphspecies_counts <- data.frame("Site Specialist" = colSums(S.S_Amphibian_1), "Mobile Generalist" = colSums(M.G_Amphibian_1), "Mobile Specialist" = colSums(M.S_Amphibian_1), "Central Forager" = colSums(C.F_Amphibian_1))
rownames(amphspecies_counts) <- "Species_Count"

#structing the df appropiately so that it is easier to use in ggplot2:
amphspecies_counts_plot <- transpose(amphspecies_counts)
rownames(amphspecies_counts_plot) <- colnames(amphspecies_counts)
colnames(amphspecies_counts_plot) <- rownames(amphspecies_counts)
amphspecies_counts_plot <- rownames_to_column(amphspecies_counts_plot, var = "Syndromes")

#barchart showing the number of Bird Species per Syndrome
amphspecies_count_per_syndrome <- ggplot(amphspecies_counts_plot, aes(x = Syndromes, y = Species_Count, fill = Syndromes)) + geom_bar(stat = "identity", color = "black") + theme_minimal() + scale_fill_manual(values=c("#f9d12e", "#00bf0b", "#fe7191", "#40dde9")) + labs(title = "Amphibian Species Count per Syndrome", y = "Number of Species") + theme(plot.title = element_text(hjust = 0.5), axis.line=element_line(color="black")) + scale_x_discrete(labels=c("Central.Forager" = "Central Forager", "Mobile.Generalist" = "Mobile Generalist", "Mobile.Specialist" = "Mobile Specialist", "Site.Specialist" = "Site Specialist")) + scale_y_continuous(expand=c(0, 0), limits=c(0, 80)) + theme(legend.position = "none") 


amphspecies_count_per_syndrome
```
